High-Efficiency, Small-Footprint Couplers Between Arbitrary Nanophotonic Waveguide Modes

### 2012-01-10 presentations/ob1_wg_paper/paper.tex
Another update on ob1 paper.

\maketitle
\begin{abstract}
We develop an algorithm for designing
    high efficiency ($\sim$95\%), 
    small-footprint (1.5-4 square vacuum wavelengths)
    couplers between arbitrary nanophotonic waveguide modes
    in two dimensions.
Our ``objective-first'' method is
    computationally fast (15 minutes on a single-core personal computer), 
    requires no trial-and-error, and
    does not require guessing a good starting design.
We demonstrate designs for various coupling problems which suggest 
    that our method allows for the design of any 
    single-mode, linear optical device.
\end{abstract}
\ocis{230.75370, 130.3990.}

@@

\begin{figure}[htb]
    \centering
    \includegraphics[width=0.8\textwidth]{1} 
    \caption{Coupler from a narrow, high-index ($\epsilon=12.25$) waveguide to
            a wide, low-index ($\epsilon=2.25$) waveguide. 
        The $H_z^\text{perfect}$ boundary values used as 
            the design objective are shown in the upper-left plots, 
            the generated structure is shown in the upper-right plot, and
            the simulated $H_z$ fields of the device are shown 
            in the bottom two plots.
        The computed efficiency of the coupler is high, $99.8\%$, and
            the device is also extremely compact, 
            convering only $36 \times 76$ grid points,
            where the vacuum wavelength is 42 grid points
            (footprint of $1.55$ square vacuum wavelengths).
        Computation time was 15 minutes on a personal computer.}
    \label{fig:fiber}
\end{figure}
\begin{figure}[htb]
    \centering
    \includegraphics[width=0.8\textwidth]{2} 
    \caption{Coupler that converts the fundamental waveguide mode to the
            second-order waveguide mode
            (a movie of the design progress is included as a
            supplementary material).
        This problem is quite difficult since the two modes are of 
            opposite symmetry.
        For example, adiabatic approaches cannot be applied to this case.
        However, our method produces a device 
            (which has the same dimensions and vacuum wavelength as 
            Fig.~\ref{fig:fiber})  
            with a coupling efficiency of $98.0\%$. 
        Computation time was 15 minutes on a personal computer.
        }
    \label{fig:mode}
\end{figure}
\begin{figure}[htb]
    \centering
    \includegraphics[width=0.8\textwidth]{3}
    \caption{Coupler between a dielectric slab waveguide to 
            an air-core waveguide.
        Here, not only are the modes of opposite symmetry,
            but the output waveguide operates on a fundamentally different
            principle (distributed reflection) than the input waveguide 
            (index guided).
        The device still achieves an efficiency of $98.9\%$, demonstrating the
            versatility of our method.
        The vacuum wavelength is 25 grid points, 
            while the device footprint is still $36 \times 76$ grid points
            (footprint of 4.38 square vacuum wavelengths).
        Computation time was 15 minutes on a personal computer.
        }
        \label{fig:aircore}
\end{figure}
\begin{figure}[htb]
    \centering
    \includegraphics[width=0.8\textwidth]{4}
    \caption{Coupler between a dielectric slab waveguide to 
            a plasmonic metal-insulator-metal waveguide.
        The efficiency of the device is $97.5\%$ and 
            has the same wavelength and footprint as the device in
            Fig.~\ref{fig:aircore}.
        Computation time was 15 minutes on a personal computer.
        }
        \label{fig:mim}
\end{figure}
\begin{figure}[htb]
    \centering
    \includegraphics[width=0.8\textwidth]{5}
    \caption{Coupler between a dielectric slab waveguide to 
            a plasmonic wire waveguide.
        The efficiency of the device is $99.1\%$ and 
            has the same wavelength and footprint as the device in
            Fig.~\ref{fig:aircore}.
        Computation time was 15 minutes on a personal computer.
        }
        \label{fig:wire}
\end{figure}

@@

This work has been supported by the 
    AFOSR MURI for Complex and Robust On-chip Nanophotonics 
    (Dr. Gernot Pomrenke), grant number FA9550-09-1-0704.
The Matlab code used is available online\cite{code}.
\clearpage
\begin{appendix}
\section{Additional mode coupler designs}
In this section, we present additional mode coupler designs.
A wide, high-index waveguide is used at both the input and output ports,
    and all permutations of couplers are fabricated among
    the first four propagating modes.
\begin{figure}[htb]
    \centering
    \includegraphics[width=0.75\textwidth]{6}
    \caption{
        Efficiency: $99.3\%$,
        footprint: $1.55$ square vacuum wavelengths.
        }
        % \label{fig:wire}
\end{figure}

\begin{figure}[htb]
    \centering
    \includegraphics[width=0.75\textwidth]{7}
    \caption{
        Efficiency: $98.3\%$,
        footprint: $1.55$ square vacuum wavelengths.
        }
        % \label{fig:wire}
\end{figure}

\begin{figure}[htb]
    \centering
    \includegraphics[width=0.75\textwidth]{8}
    \caption{
        Efficiency: $90.6\%$,
        footprint: $1.55$ square vacuum wavelengths.
        }
        % \label{fig:wire}
\end{figure}

\begin{figure}[htb]
    \centering
    \includegraphics[width=0.75\textwidth]{9}
    \caption{
        Efficiency: $96.8\%$,
        footprint: $1.55$ square vacuum wavelengths.
        }
        % \label{fig:wire}
\end{figure}

\begin{figure}[htb]
    \centering
    \includegraphics[width=0.75\textwidth]{10}
    \caption{
        Efficiency: $86.3\%$,
        footprint: $1.55$ square vacuum wavelengths.
        }
        % \label{fig:wire}
\end{figure}

\begin{figure}[htb]
    \centering
    \includegraphics[width=0.75\textwidth]{11}
    \caption{
        Efficiency: $80.1\%$,
        footprint: $1.55$ square vacuum wavelengths.
        }
        % \label{fig:wire}
\end{figure}
\clearpage
\section{Additional designs with wide, low-index input waveguide}
\begin{figure}[htb]
    \centering
    \includegraphics[width=0.75\textwidth]{12}
    \caption{
        Efficiency: $96.9\%$,
        footprint: $1.55$ square vacuum wavelengths.
        }
        % \label{fig:wire}
\end{figure}

\begin{figure}[htb]
    \centering
    \includegraphics[width=0.75\textwidth]{13}
    \caption{
        Efficiency: $99.0\%$,
        footprint: $4.38$ square vacuum wavelengths.
        }
        % \label{fig:wire}
\end{figure}

\begin{figure}[htb]
    \centering
    \includegraphics[width=0.75\textwidth]{14}
    \caption{
        Efficiency: $96.7\%$,
        footprint: $4.38$ square vacuum wavelengths.
        }
        % \label{fig:wire}
\end{figure}

\begin{figure}[htb]
    \centering
    \includegraphics[width=0.75\textwidth]{15}
    \caption{
        Efficiency: $99.7\%$,
        footprint: $4.38$ square vacuum wavelengths.
        }
        % \label{fig:wire}
\end{figure}
\clearpage

\section{Additional designs with multiple output plasmonic waveguides}
\begin{figure}[htb]
    \centering
    \includegraphics[width=0.75\textwidth]{16}
    \caption{
        Efficiency: $94.0\%$,
        footprint: $4.38$ square vacuum wavelengths.
        }
        % \label{fig:wire}
\end{figure}

\begin{figure}[htb]
    \centering
    \includegraphics[width=0.75\textwidth]{17}
    \caption{
        Efficiency: $97.3\%$,
        footprint: $4.38$ square vacuum wavelengths.
        }
        % \label{fig:wire}
\end{figure}

\begin{figure}[htb]
    \centering
    \includegraphics[width=0.75\textwidth]{18}
    \caption{
        Efficiency: $98.5\%$,
        footprint: $4.38$ square vacuum wavelengths.
        }
        % \label{fig:wire}
\end{figure}

\begin{figure}[htb]
    \centering
    \includegraphics[width=0.75\textwidth]{19}
    \caption{
        Efficiency: $97.7\%$,
        footprint: $4.38$ square vacuum wavelengths.
        }
        % \label{fig:wire}
\end{figure}

\begin{figure}[htb]
    \centering
    \includegraphics[width=0.75\textwidth]{20}
    \caption{
        Efficiency: $95.6\%$,
        footprint: $4.38$ square vacuum wavelengths.
        }
        % \label{fig:wire}
\end{figure}

\begin{figure}[htb]
    \centering
    \includegraphics[width=0.75\textwidth]{21}
    \caption{
        Efficiency: $87.5\%$,
        footprint: $4.38$ square vacuum wavelengths.
        }
        % \label{fig:wire}
\end{figure}
\clearpage
\end{appendix}
\end{document}

>>>>

\maketitle
\begin{abstract}
We develop an algorithm for designing
    high efficiency ($\sim$98\%), 
    small-footprint (1.5-4 square vacuum wavelengths)
    couplers between arbitrary nanophotonic waveguide modes
    in two dimensions.
Our ``objective-first'' method is
    computationally fast (15 minutes on a single-core personal computer), 
    requires no trial-and-error, and
    does not require guessing a good starting design.
We demonstrate designs for various coupling problems which suggest 
    that our method allows for the design of any 
    single-mode, linear optical device.
\end{abstract}
\ocis{230.75370, 130.3990.}

@@

\begin{figure}[htb]
    \centering
    \includegraphics[width=0.8\textwidth]{1} 
    \caption{Coupler from a narrow, high-index ($\epsilon=12.25$) waveguide to
            a wide, low-index ($\epsilon=2.25$) waveguide. 
        The $H_z^\text{perfect}$ boundary values used as 
            the design objective are shown in the upper-left plots, 
            the generated structure is shown in the upper-right plot, and
            the simulated $H_z$ fields of the device are shown 
            in the bottom two plots.
        The computed efficiency of the coupler is high, $99.8\%$, and
            the device is also extremely compact, 
            convering only $36 \times 76$ grid points,
            where the vacuum wavelength is 42 grid points
            (footprint of $1.55$ square vacuum wavelengths).
        Computation time was 15 minutes on a personal computer.}
    \label{fig:fiber}
\end{figure}
\begin{figure}[htb]
    \centering
    \includegraphics[width=0.8\textwidth]{2} 
    \caption{Coupler that converts the fundamental waveguide mode to the
            second-order waveguide mode
            (a movie of the design progress is included as a
            supplementary material).
        This problem is quite difficult since the two modes are of 
            opposite symmetry.
        For example, adiabatic approaches cannot be applied to this case.
        However, our method produces a device 
            (which has the same dimensions and vacuum wavelength as 
            Fig.~\ref{fig:fiber})  
            with a coupling efficiency of $98.0\%$. 
        Computation time was 15 minutes on a personal computer.
        }
    \label{fig:mode}
\end{figure}
\begin{figure}[htb]
    \centering
    \includegraphics[width=0.8\textwidth]{3}
    \caption{Coupler between a dielectric slab waveguide to 
            an air-core waveguide.
        Here, not only are the modes of opposite symmetry,
            but the output waveguide operates on a fundamentally different
            principle (distributed reflection) than the input waveguide 
            (index guided).
        The device still achieves an efficiency of $98.9\%$, demonstrating the
            versatility of our method.
        The vacuum wavelength is 25 grid points, 
            while the device footprint is still $36 \times 76$ grid points
            (footprint of 4.38 square vacuum wavelengths).
        Computation time was 15 minutes on a personal computer.
        }
        \label{fig:aircore}
\end{figure}
\begin{figure}[htb]
    \centering
    \includegraphics[width=0.8\textwidth]{4}
    \caption{Coupler between a dielectric slab waveguide to 
            a plasmonic metal-insulator-metal waveguide.
        The efficiency of the device is $97.5\%$ and 
            has the same wavelength and footprint as the device in
            Fig.~\ref{fig:aircore}.
        Computation time was 15 minutes on a personal computer.
        }
        \label{fig:mim}
\end{figure}
\clearpage
\begin{figure}[htb]
    \centering
    \includegraphics[width=0.8\textwidth]{5}
    \caption{Coupler between a dielectric slab waveguide to 
            a plasmonic wire waveguide.
        The efficiency of the device is $99.1\%$ and 
            has the same wavelength and footprint as the device in
            Fig.~\ref{fig:aircore}.
        Computation time was 15 minutes on a personal computer.
        }
        \label{fig:wire}
\end{figure}

@@

This work has been supported by the 
    AFOSR MURI for Complex and Robust On-chip Nanophotonics 
    (Dr. Gernot Pomrenke), grant number FA9550-09-1-0704.
The Matlab code used is available online\cite{code}.
\clearpage
\begin{appendix}
\section{Additional mode coupler designs}
In this section, we present additional mode coupler designs.
A wide, high-index waveguide is used at both the input and output ports,
    and all permutations of couplers are fabricated among
    the first four propagating modes.
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\textwidth]{6}
    \caption{
        Coupler from the first-order to the second-order mode 
            of a wide dielectric waveguide.
        Efficiency: $99.3\%$,
        footprint: $1.55$ square vacuum wavelengths.
        }
        % \label{fig:wire}
\end{figure}
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\textwidth]{7}
    \caption{
        Coupler from the first-order to the third-order mode 
            of a wide dielectric waveguide.
        Efficiency: $98.3\%$,
        footprint: $1.55$ square vacuum wavelengths.
        }
        % \label{fig:wire}
\end{figure}
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\textwidth]{8}
    \caption{
        Coupler from the first-order to the fourth-order mode 
            of a wide dielectric waveguide.
        Efficiency: $90.6\%$,
        footprint: $1.55$ square vacuum wavelengths.
        }
        % \label{fig:wire}
\end{figure}
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\textwidth]{9}
    \caption{
        Coupler from the second-order to the third-order mode 
            of a wide dielectric waveguide.
        Efficiency: $96.8\%$,
        footprint: $1.55$ square vacuum wavelengths.
        }
        % \label{fig:wire}
\end{figure}
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\textwidth]{10}
    \caption{
        Coupler from the second-order to the fourth-order mode 
            of a wide dielectric waveguide.
        Efficiency: $86.3\%$,
        footprint: $1.55$ square vacuum wavelengths.
        }
        % \label{fig:wire}
\end{figure}
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\textwidth]{11}
    \caption{
        Coupler from the third-order to the fourth-order mode 
            of a wide dielectric waveguide.
        Efficiency: $80.1\%$,
        footprint: $1.55$ square vacuum wavelengths.
        }
        % \label{fig:wire}
\end{figure}
\clearpage
\section{Additional designs with wide, low-index input waveguide}
We now reproduce Figs~\ref{fig:mode}-\ref{fig:wire}
    but instead use a wide, low-index waveguide as the input.
Once again, this is to demonstrate the generality of our method;
    namely, that it can be applied to the design of nearly any
    single-mode, linear nanophotonic device.
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\textwidth]{12}
    \caption{
        Coupler from a wide, low-index waveguide to the
            second-order mode of a narrow, high-index waveguide.
        Efficiency: $96.9\%$,
        footprint: $1.55$ square vacuum wavelengths.
        }
        % \label{fig:wire}
\end{figure}
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\textwidth]{13}
    \caption{
        Coupler from a wide, low-index waveguide to an 
            ``air-core'' waveguide mode.
        Efficiency: $99.0\%$,
        footprint: $4.38$ square vacuum wavelengths.
        }
        % \label{fig:wire}
\end{figure}
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\textwidth]{14}
    \caption{
        Coupler from a wide, low-index waveguide to a
            metal-insulator-metal plasmonic waveguide mode.
        Efficiency: $96.7\%$,
        footprint: $4.38$ square vacuum wavelengths.
        }
        % \label{fig:wire}
\end{figure}
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\textwidth]{15}
    \caption{
        Coupler from a wide, low-index waveguide to a
            plasmonic wire waveguide mode.
        Efficiency: $99.7\%$,
        footprint: $4.38$ square vacuum wavelengths.
        }
        % \label{fig:wire}
\end{figure}
\clearpage

\section{Additional designs with multiple output plasmonic waveguides}
We also present ``selector'' designs where photons are coupled to 
    only one of five possible plasmonic output waveguides.
We demonstrate these selector designs for both metal-insulator-metal and
    metal-wire plasmonic waveguides.
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\textwidth]{16}
    \caption{
        Coupler from a dielectric waveguide to the 
            lowest branch of a set of five plasmonic wire waveguides.
        Efficiency: $94.0\%$,
        footprint: $4.38$ square vacuum wavelengths.
        }
        % \label{fig:wire}
\end{figure}
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\textwidth]{17}
    \caption{
        Coupler from a dielectric waveguide to the 
            second branch of a set of five plasmonic wire waveguides.
        Efficiency: $97.3\%$,
        footprint: $4.38$ square vacuum wavelengths.
        }
        % \label{fig:wire}
\end{figure}
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\textwidth]{18}
    \caption{
        Coupler from a dielectric waveguide to the 
            middle branch of a set of five plasmonic wire waveguides.
        Efficiency: $98.5\%$,
        footprint: $4.38$ square vacuum wavelengths.
        }
        % \label{fig:wire}
\end{figure}
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\textwidth]{19}
    \caption{
        Coupler from a dielectric waveguide to the 
            middle branch of a set of five plasmonic 
            metal-insulator-metal waveguides.
        Efficiency: $97.7\%$,
        footprint: $4.38$ square vacuum wavelengths.
        }
        % \label{fig:wire}
\end{figure}
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\textwidth]{20}
    \caption{
        Coupler from a dielectric waveguide to the 
            fourth branch of a set of five plasmonic 
            metal-insulator-metal waveguides.
        Efficiency: $95.6\%$,
        footprint: $4.38$ square vacuum wavelengths.
        }
        % \label{fig:wire}
\end{figure}
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\textwidth]{21}
    \caption{
        Coupler from a dielectric waveguide to the 
            uppermost branch of a set of five plasmonic 
            metal-insulator-metal waveguides.
        Efficiency: $87.5\%$,
        footprint: $4.38$ square vacuum wavelengths.
        }
        % \label{fig:wire}
\end{figure}
\clearpage
\end{appendix}
\end{document}
